---
title: "STA 483 Project Part 3"
author: "Brian Lambert"
date: "12/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/brianlambert/Desktop/STA_483/project_pt2")
library(tidyverse)
library(lubridate)
library(forecast)
library(knitr)
library(dplyr)
library(doBy)
library(gridExtra)
library(WeightedPortTest)
library(dummies)
```

```{r}
years = c("2001","2002","2003","2004","2005","2006","2007","2008","2009","2010","2011","2012","2013","2014","2015","2016","2017")
data = data.frame(matrix(ncol = 4, nrow = 0))
for (year in years) {
  path = paste("madrid_01_17/madrid_", year, ".csv", sep="")
  new_data = read.csv(path)
  new_data = new_data[,c("date", "CO", "NO_2", "station")]
  data = rbind(data, new_data)
}

data = data %>%
  mutate(id = as.factor(station))

stations = read.csv("stations.csv")
stations = stations %>%
  mutate(id = as.factor(id))

data = left_join(data, stations, by = "id")
data = data %>% 
  drop_na(name)

head(data)
```


```{r}
ym_station_data = data %>%
  mutate(Month = month(date)) %>%
  mutate(Year = year(date)) %>% 
  group_by(id, Year, Month) %>%
  summarise_at(c("CO", "NO_2"), mean, na.rm = TRUE)

head(ym_station_data)
tail(ym_station_data)
```


```{r}
stations_to_keep = c("28079004", "28079008", "28079016", "28079024", "28079036", "28079039")

data_ym  = data %>%
  filter(id %in% stations_to_keep) %>%
  mutate(Month = month(date)) %>%
  mutate(Year = year(date)) %>%
  group_by(Year, Month) %>%
  filter(Year != "2018") %>%
  summarise_at(c("CO", "NO_2"), mean, na.rm = TRUE)
```

### Model building
+ sArima

```{r}
Year.Month.specific.ts <- ts(data_ym$CO, start=c(2001,1), freq=12)

autoplot(Year.Month.specific.ts)
## definite trend here; not stationary; dont really see anything with increasing/decreasing variance

autoplot(log(Year.Month.specific.ts))
## this definitely fixes the stationarity
```

Thus, we will be working with the log of the time series. Now to determine what model to fit. 

```{r}

grid.arrange(ggAcf(log(Year.Month.specific.ts), lag=36),
             ggPacf(log(Year.Month.specific.ts), lag=36),
             nrow=1)

```


Definitely some sARIMA process here. On the ACF there is significance at lag 12, 24, 36 etc. PACF seems to either cut at 2 or decay. Thus, seasonal component of AR(2) or an ARMA(1,2) or an

```{r}

seasonal.mod1 <- Arima(Year.Month.specific.ts, seasonal=c(1,0,2), lambda=0)
seasonal.mod2 <- Arima(Year.Month.specific.ts, seasonal=c(2,0,2), lambda = 0)
seasonal.mod1
seasonal.mod2
## AIC and BIC suggest that the sesaonal component that will perform better will be an SARIMA (2,0,2)

grid.arrange(ggAcf(seasonal.mod2$residuals),
             ggPacf(seasonal.mod2$residuals),
             nrow=1)

```

Fixes most of the issues. ACF could decay or cut, definite PACF cut. ACF would cut at 2, PACF at 1 MA(1)? ARMA (1,1)? AR(2) Let's model


```{r}

Arima(Year.Month.specific.ts, order=c(1,0,1), seasonal=c(2,0,2), lambda=0)
Arima(Year.Month.specific.ts, order=c(0,0,1), seasonal=c(2,0,2), lambda=0) 
Arima(Year.Month.specific.ts, order=c(2,0,0), seasonal=c(2,0,2), lambda=0) 

Weighted.Box.test(Arima(Year.Month.specific.ts, order=c(0,1,1), seasonal=c(1,1,1))$residuals, type = "Ljung")

auto.arima(Year.Month.specific.ts)

# auto.arima: ARIMA(1,1,1)(2,1,1)[12]
# auto.arima (stepwise=F, approximation=F): ARIMA(0,1,1)(1,1,1)[12]

```

Going to keep both the sARIMA(2,0,0)(2,0,2) as well as the sARIMA(1,0,1)(2,0,2) to look at their residuals

```{r}

sarima.specific1 <- Arima(Year.Month.specific.ts, order=c(1,0,1), seasonal=c(2,0,2), lambda=0)
grid.arrange(ggAcf(sarima.specific1$residuals),
             ggPacf(sarima.specific1$residuals),
             nrow=1)

#sarima.specific2 <- Arima(Year.Month.specific.ts, order=c(2,0,1), seasonal=c(1,1,1))
sarima.specific2 <- Arima(Year.Month.specific.ts, order=c(2,0,0), seasonal=c(2,0,2))
grid.arrange(ggAcf(sarima.specific2$residuals),
             ggPacf(sarima.specific2$residuals),
             nrow=1)
```

only the sARIMA(2,0,0)(2,0,2) seems to have reasonably white noise residuals. Thus, I will keep it for the final validation study.


# CV
```{r}
ts.cv.study <- function(time.pt) {
  
  # gets index of specific time pt passed to function
  time.spot <- which(abs(time(Year.Month.specific.ts)-time.pt)<0.001)

  unemp.test <- window(Year.Month.specific.ts, start=time(Year.Month.specific.ts)[time.spot+1], end=time(Year.Month.specific.ts)[time.spot+5])
  unemp.train <- window(Year.Month.specific.ts, start = c(2008,1), end=time.pt)
  
  SIN <- COS <- matrix(nrow=length(unemp.train), ncol=2)
  for(i in 1:2) {
    SIN[,i] <- sin(2*pi*i*time(unemp.train))
    COS[,i] <- cos(2*pi*i*time(unemp.train))
  }
  
  deterministic.fit <- Arima(unemp.train, order=c(2,0,0), xreg=cbind(SIN,COS), method = "CSS")
  hw.fit <- HoltWinters(unemp.train, seasonal = "multiplicative", beta = FALSE)
  arima.fit <- Arima(unemp.train, order=c(2,0,0), seasonal=c(2,0,2), lambda = 0, method = "CSS")
  
  fake.data <- ts(rep(NA, 48), start=c(2018,1), freq=12)
  SIN.fut <- cos.fut <- matrix(nrow=length(fake.data), ncol=2) 
  COS.fut <- cos.fut <- matrix(nrow=length(fake.data), ncol=2) 
  for(i in 1:2) {
    SIN.fut[,i] <- sin(2*pi*i*time(fake.data))
    COS.fut[,i] <- cos(2*pi*i*time(fake.data))
  }

  deterministic.pred <- forecast(deterministic.fit, xreg=cbind(SIN.fut, COS.fut), h=length(unemp.test))
  hw.pred <- forecast(hw.fit, h=length(unemp.test))
  arima.pred <- forecast(arima.fit, h=length(unemp.test))

 c(accuracy(deterministic.pred, unemp.test)[2,3],
    accuracy(hw.pred, unemp.test)[2,3],
    accuracy(arima.pred, unemp.test)[2,3],
    accuracy(deterministic.pred, unemp.test)[2,2],
    accuracy(hw.pred, unemp.test)[2,2],
    accuracy(arima.pred, unemp.test)[2,2]) 
}
```

```{r}
januarys = c(2013,2014,2015,2016,2017)
cv.output <- sapply(januarys, FUN=ts.cv.study)

my.dates <- time(window(Year.Month.specific.ts, start=c(2012,3), end=c(2017, 7)))
cv.output <- sapply(as.vector(my.dates), FUN=ts.cv.study)
```

```{r}
mae.tab <- t(cv.output[1:3,])   # Flip rows and columns
rmse.tab <- t(cv.output[4:6,])
colnames(mae.tab) <- colnames(rmse.tab) <- c("Deterministic", "HW, no trend", "ARIMA(2,0,0)(2,0,2)")

mae.summary <- cbind(apply(mae.tab, 2, mean), apply(mae.tab, 2, sd), apply(mae.tab, 2, median) )
rmse.summary <- cbind(apply(rmse.tab, 2, mean), apply(rmse.tab, 2, sd), apply(rmse.tab, 2, median) )
colnames(mae.summary) <- colnames(rmse.summary) <- c("Mean", "SD", "Median")

kable(mae.summary)
kable(rmse.summary)
```

```{r}
final.fit1 = Arima(Year.Month.specific.ts, order=c(2,0,1), seasonal = c(1,1,1))
final.fit2 = Arima(Year.Month.specific.ts, order=c(1,1,1), seasonal = c(2,1,1))
final.fit3 = Arima(Year.Month.specific.ts, order=c(0,1,1), seasonal = c(1,1,1))
final.pred1 <- forecast(final.fit1, h=5)
final.pred2 <- forecast(final.fit2, h=5)
final.pred3 <- forecast(final.fit3, h=5)

autoplot(window(Year.Month.specific.ts, start=c(2015,1)) ) +
  autolayer(final.pred1, PI=FALSE, series="ARIMA(2,0,1)(1,1,1)") +
  autolayer(final.pred2, PI=FALSE, series="ARIMA(1,1,1)(2,1,1)") +
  autolayer(final.pred3, PI=FALSE, series="ARIMA(0,1,1)(1,1,1)") 
```


# Problem 2

```{r}
stations_to_keep2 = c("28079004","28079036","28079038","28079039","28079040","28079008",
                      "28079011","28079016","28079017","28079018","28079024","28079027","28079035")

data_hourly = data %>%
  filter(id %in% stations_to_keep2) %>%
  mutate(Year = year(date),
         Month= month(date),
         Day = day(date),
         Hour = hour(date),
         Weekday = wday(date)) %>%
  group_by(Year, Month, Day, Hour, Weekday) %>%
  summarise_at(c("NO_2"), mean, na.rm = TRUE)
```

## half of data
```{r}
hourly.half = data_hourly %>%
  filter(Month %in% c(1,12) & Year > 2016)
```

## just Saturday data
```{r}
hourly.sat = data_hourly %>%
  filter(Weekday == 7)

hourly.sat.ts = ts(hourly.sat$NO_2, start=c(2009,1,1,0), freq=24)
autoplot(hourly.sat.ts)

grid.arrange(ggAcf(hourly.ts, lag=48),
             ggPacf(hourly.ts, lag=48),
             nrow=1)

hourly.sat.model1 <- Arima(hourly.sat.ts, seasonal=c(1,0,0), method="CSS")
hourly.sat.model2 <- Arima(hourly.sat.ts, seasonal=c(2,0,0), method="CSS")
hourly.sat.model3 <- Arima(hourly.sat.ts, seasonal=c(2,1,1), method="CSS")
hourly.sat.model1
hourly.sat.model2
hourly.sat.model3

##thus we will use seasonal component (2,0,1) because it has the highest log likelihood

grid.arrange(ggAcf(hourly.sat.model2$residuals, lag=48),
             ggPacf(hourly.sat.model2$residuals, lag=48),
             nrow=1)

##ACF decay at 24, PACF decay or cut. AR(1), AR(2) ARMA(1,1), ARMA(2,1), MA(1)

hourly.sat.model.full1 <- Arima(hourly.sat.ts, seasonal=c(2,0,1), order=c(1,0,0), method="CSS")
hourly.sat.model.full2 <- Arima(hourly.sat.ts, seasonal=c(2,0,1), order=c(2,0,0), method="CSS")
hourly.sat.model.full3 <- Arima(hourly.sat.ts, seasonal=c(2,0,1), order=c(1,0,2), method="CSS")
hourly.sat.model.full4 <- Arima(hourly.sat.ts, seasonal=c(2,0,2), order=c(2,0,3), method="CSS")
hourly.sat.model.full5 <- Arima(hourly.sat.ts, seasonal=c(2,0,1), order=c(0,0,1), method="CSS")

auto.arima(hourly.sat.ts)

hourly.sat.model.full6 <-Arima(hourly.sat.ts, seasonal=c(2,1,2), order=c(0,0,2), method="CSS")

grid.arrange(ggAcf(hourly.sat.model.full6$residuals, lag=48),
             ggPacf(hourly.sat.model.full6$residuals, lag=48),
             nrow=1)

```

## The last saturday
```{r}
head(hourly.sat)
 
hour.no2 = hourly.sat[c("Hour","NO_2")]
hour.no2 = hour.no2[c(25:nrow(hour.no2)), ]

previous.hours = spread(hourly.sat, "Hour", "NO_2")[c("0","1","2","3","4","5","6","7",
                                                      "8","9","10","11","12","13","14","15",
                                                      "16","17","18","19","20","21","22","23")]

full.repeated = NULL
for (i in 2:887) {
  
  temp.repeated = NULL
  previous.day = previous.hours[i-1,]
    
  for (j in 1:24) {
    temp.repeated = rbind(temp.repeated, previous.day)
  }
  
  full.repeated = rbind(full.repeated, temp.repeated)
}

final.data = as.data.frame(cbind(full.repeated, hour.no2))  

final.data = dummy.data.frame(final.data,names=c("Hour"),sep="_")

fit1 = lm(NO_2 ~ ., data = final.data)
summary(fit1)
```



```{r}
for (i in 0:23) {
  new.data = c(48.076923,37.307692,30.230769,23.153846,18.692308,15.153846,
            14.076923,14.230769,20.000000,22.230769,25.769231,27.230769,
            27.461538,25.461538,23.076923,20.307692,18.769231,21.076923,
            30.153846,47.461538,67.692308,78.153846,82.307692,83.461538)
  
  new.dummies = c(rep(0,24))
  new.dummies[i+1] = 1
  
  new.data = append(new.data, new.dummies)

  new.df = as.data.frame(t(as.data.frame(new.data)))
  colnames(new.df) = c("0","1","2","3","4","5","6","7","8","9","10","11","12","13","14","15",
                     "16","17","18","19","20","21","22","23","Hour_0","Hour_1","Hour_2","Hour_3"
                     ,"Hour_4","Hour_5","Hour_6","Hour_7","Hour_8","Hour_9","Hour_10","Hour_11"
                     ,"Hour_12","Hour_13","Hour_14","Hour_15","Hour_16","Hour_17","Hour_18","Hour_19"
                     ,"Hour_20","Hour_21","Hour_22","Hour_23")

  print(i)
  print(suppressWarnings(predict(fit1, new.df)))
}

```

```{r}
library(DAAG)
cv.lm(data = final.data, fit1, m=5, plotit=TRUE, printit=TRUE)
```

## Average of last 4 Saturdays
```{r}
sat.avgs = hourly.sat[c((nrow(hourly.sat)-(24*4)+1):nrow(hourly.sat)), c("Hour","NO_2")]
sat.avgs =  sat.avgs %>%
  group_by(Hour) %>%
  summarise(no2.mean = mean(NO_2))
```

#########################################

```{r}
hourly.ts = ts(hourly.half$NO_2, start = c(2017,1,1,0), frequency = 1488)
autoplot(hourly.ts)

```

We used frequency = 1408 becuase we excluded all months except November, December, and January and there are 2208 total hours in those three months. Moreover, this function looks stationary, however seasonal.


```{r}

grid.arrange(ggAcf(hourly.ts, lag=48),
             ggPacf(hourly.ts, lag=48),
             nrow=1)

hourly.ts.model <- Arima(hourly.ts, order=c(2,0,1), method = "CSS")

hourly.ts.model.full <- Arima(hourly.ts, seasonal=c(1,0,0), order=c(1,0,1), method="CSS")

grid.arrange(ggAcf(hourly.ts.model$residuals, lag=48),
             ggPacf(hourly.ts.model$residuals, lag=48),
             nrow=1)

```


