---
title: "STA 483 Project Part 1"
author: "Brian Lambert"
date: "October 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(forecast)
library(knitr)
library(gridExtra)
```
# 1

### load and clean data 

```{r}
# madrid = read.csv("C:/Users/Brian Lambert/Desktop/Miami_Fall_2018/STA_483/madrid_2017.csv")
madrid = read.csv("/Users/brianlambert/Desktop/STA_483/madrid_2017.csv")
#stations = read.csv("C:/Users/Brian Lambert/Desktop/Miami_Fall_2018/STA_483/stations.csv")
stations = read.csv("/Users/brianlambert/Desktop/STA_483/stations.csv")

no2 = madrid[!is.na(madrid$NO_2),] # remove all rows with NO_2 = NA
no2.pde = no2[no2$station == 28079004,] # keep only recordings from Pza. De Espana (station #28079004)
no2.pde = no2.pde %>%
  select(date, NO_2) # remove all unnecessary columns
```

# 2

### plot time series and describe / discuss

```{r}
pde.ts = ts(no2.pde$NO_2, start=c(2017,6,1,1), frequency = 8760)

autoplot(pde.ts, scales = "%Y-%m") + 
  labs(x="Time", y="NO2 Levels (ug/m^3)") +
  ggtitle("NO2 Levels at the Pza. De Espana Recording Station") +
  theme_minimal()
```

The series appears to be following some sort of sine or cosine wave pattern. I believe this time series is clearly non-stationary becuase the mean is visually changing throughout the year, as well as the variance. Also, as discussed in lab 8, there appears to be large spikes in the variance coinciding with mean increases which are then followed by a period of decreased mean and variance. All of these descriptions must be taken with a grain of salt, however, because the only available data was strictly collected from June, 2017 to January, 2018 so we cannot be sure that these patterns are constant year-to-year or were even a continuation of patterns in the first 5 months of 2017. 

# 3

### Seasonality of NO2

```{r}

no2 = madrid[!is.na(madrid$NO_2),] # remove all rows with CO = NA
no2 = no2 %>%
  select(date, NO_2, station) # remove all unnecessary columns

# use POSIXlt to extract month
dates.no2 = as.POSIXlt(no2$date, format = "%Y-%m-%d%H:%M")
no2.month = data.frame("NO_2" = no2$NO_2, "station" = no2$station)
no2.month$year = dates.no2$year + 1900
no2.month$month = dates.no2$mon + 1

no2.month.mean = no2.month %>%
  group_by(month) %>%
  summarise(mean = mean(NO_2),
            sd = sd(NO_2))
no2.month.mean
```

### Monthly Mean NO2 Plot

```{r}
ggplot(no2.month.mean, aes(month, mean)) +
  geom_line(size = 2) +
  ggtitle("Montly Mean NO2 Levels for all of Madrid")
```

### Monthly NO2 Standard Deviation Plot

```{r}
ggplot(no2.month.mean, aes(month, sd)) +
  geom_line(size = 2) +
  ggtitle("Montly NO2 Standard Deviation for all of Madrid")
```

The above table and plots make it clear that this data is highly seasonal. We clearly see that not only the mean, but the standard devaiatn vary month-to-month, following a bi-yearly sine or cosine wave pattern.

# 4

### Monthly mean CO levels by station table

```{r}
co = madrid[!is.na(madrid$CO),] # remove all rows with CO = NA
co = co %>%
  select(date, CO, station) # remove all unnecessary columns

# use POSIXlt to extract month
dates.co = as.POSIXlt(co$date, format = "%Y-%m-%d%H:%M")
co.mon = data.frame("CO" = co$CO, "station" = co$station)
co.mon$month = dates.co$mon + 1

# grouped by station then month
means.co = co.mon %>%
  group_by(station, month) %>%
  summarise(mean = mean(CO))
head(means.co)
```

### Which site appears to have the largest levels of carbon monoxide?

```{r}
means.co$station = as.factor(means.co$station) # convert station to factor for plotting
means.co$month = as.factor(means.co$month) # convert month to factor for plotting

ggplot(means.co, aes(month, mean)) +
  labs(y="mean CO Levels (mg/m^3)") +
  geom_line(aes(group = station, color = station), size = 2) +
  ggtitle("Montly Mean CO Levels by Station")
```

It appears that for the first half of the year the station with the highest monthly mean CO levels is Farolillo (28079018) and for the second half of the year it's Pza. de Espana (28079004).

### Which month appears to have the largest levels of carbon monoxide?

```{r}
# grouped by month only
means.co.mon = co.mon %>%
  group_by(month) %>%
  summarise(mean = mean(CO))

# month with the highest average CO levels
means.co.mon[which.max(means.co.mon$mean),]
```

The month with the largest levels of carbon monoxide levels is January.

# 5

### monthly CO standard deviations by station table

```{r}
# grouped by station then month
sd.co = co.mon %>%
  group_by(station, month) %>%
  summarise(sd = sd(CO))
head(sd.co)
```

### Which site appears to be most variable in terms of carbon monoxide?

```{r}
sd.co$station = as.factor(sd.co$station) # convert station to factor for plotting
sd.co$month = as.factor(sd.co$month) # convert month to factor for plotting

ggplot(sd.co, aes(month, sd)) +
  geom_line(aes(group = station, color = station), size = 2) +
  ggtitle("Monthly CO Standard Deviation Levels by Station")
```

The most variable station in terms of CO is Barrio del Pilar (28079039), while the least variable station in terms of CO is Casa de Campo (28079024).

# 6

### Hourly mean NO2 levels by station table (only top 5 yearly stations)

```{r}
no2 = madrid[!is.na(madrid$NO_2),] # remove all rows with CO = NA
no2 = no2 %>%
  select(date, NO_2, station) # remove all unnecessary columns

# use POSIXlt to extract month
dates.no2 = as.POSIXlt(no2$date, format = "%Y-%m-%d%H:%M")
no2.hour = data.frame("NO_2" = no2$NO_2, "station" = no2$station)
no2.hour$hour = dates.no2$hour
no2.hour$year = dates.no2$year + 1900

# grouped by station then year
means.no2.year = no2.hour %>%
  group_by(station, year) %>%
  summarise(mean = mean(NO_2)) %>%
  arrange(desc(mean))
head(means.no2.year, n=5) # top 5 stations to display in hourly table

# grouped by station then hour
means.no2 = no2.hour %>%
  group_by(station, hour) %>%
  summarise(mean = mean(NO_2)) %>%
  filter(station %in% c(28079008, 28079056, 28079004, 28079035, 28079017))
head(means.no2)
```

The above table appears to be broken into two groups: the first are the stations with the 2 highest mean NO2 levels, each within the (59-62) range, and the next are the next three highest, each falling into a (48.6 - 49.4) range. There is a clear distinction between the top two stations mean NO2 levels and that of the rest, so we can likely find some factor shared between those 2 stations that causes them to have such significantly higher mean levels than the other 3. The elevation is fairly different between these two top sites so we can probably rule that out as a significant predictive factor of NO2 level.

# 7

### Time Series plot of the mean NO2 levels for all of madrid

```{r}
no2.hour$month = dates.no2$mon + 1
no2.hour$day = dates.no2$wday
head(no2.hour)

means.no2.all = no2.hour %>% 
  group_by(year, month, day, hour) %>% 
  summarise(mean = mean(NO_2))
head(means.no2.all)

no2.all.ts = ts(means.no2.all$mean, start=c(2017,1,0,0), frequency = 8760)
autoplot(no2.all.ts, bands = T) +
  labs(x="Time", y="NO2 Levels (ug/m^3)") +
  ggtitle("Hourly Mean NO2 Levels for all of Madrid") +
  theme_minimal()
```

The above time series plot again shows a sine or cosine type wave pattern like before, but this time we can clearly see an even larger scale wave pattern that appears to follow an increasing parabolic pattern for the last six months of the year before reaching it's peak in December and once again dropping in January. It is reasonable to belive that this increasing parabolic pattern will repeat from January - June. Thus, we have multiple clear patterns for mean NO2 levels in Madrid throughout the year.

#8 

### Map of Mean NO2 Levels at Each Station

```{r}
means.no2.map = no2.hour %>%
  group_by(station) %>%
  summarise(mean = mean(NO_2))
colnames(means.no2.map)[1] = "id" # change variable name to match "stations" data so we can join 

means.no2.map = means.no2.map %>%
  left_join(stations, by = "id")

##########
# retrieved from stackoverflow. I set up a Google Static Maps API and created a key but it would throw an error when retrieving.
# The comment on the post said there's a issue with the currently published version and provided the following lines to fix ..
if(!requireNamespace("devtools")) install.packages("devtools")
devtools::install_github("dkahle/ggmap", ref = "tidyup")
library(ggmap)
register_google(key = "AIzaSyB5C3GHAmzoiwLbt8tSEyxmrx7OEnw0zwY") 
# citation:
# D. Kahle and H. Wickham. ggmap: Spatial Visualization with ggplot2. The R Journal, 5(1), 144-161. URL
#  http://journal.r-project.org/archive/2013-1/kahle-wickham.pdf
##########

spain<- get_googlemap(center = c(lon = -3.700279, lat = 40.396375), zoom=11,maptype = "roadmap", color = "bw") 
# used google maps to find general latitude and longitude of Madrid

ggmap(spain) +
  ggtitle("Mean NO2 Levels by Station in Madrid") +
  geom_point(data=means.no2.map, 
             aes(x=lon, y=lat, size=2, colour = mean), alpha=I(0.7)) +
  scale_colour_gradient(low = "red", high = "green")
```

